Infinite Diameter Confidence Sets in Hedges' Publication Bias Model

### 2020-08-15 infiniteci.tex
Updates from Overleaf

A meta-analysis is a statistical analysis that quantitatively combines
results from separate scientific studies. When the studies measure
the same phenomenon, pooling of information allows us to predict the
common effect size with larger precision than we could have done with
one study alone. Meta-analyses are ubiquitous in the empirical sciences
and forms a key component of most systematic reviews such as Cochrane's
reviews \citep{Higgins2019-vv}.

Most meta-analytic techniques assume honest and unbiased reporting
of results. But there is ample evidence that the scientific literature
is not unbiased, as studies with\emph{ }significant \emph{p}-values
tend to be published with a greater probability than other studies
\citep{Easterbrook1991-ph}, a phenomenon called publication bias
by \citet{sterling1959publication} and the file-drawer problem by
\citet{Rosenthal1979-pm}. When publication bias is present, there
is no reason to trust the results of meta-analytic methods that do
not account for it, as the parameter estimates will be inconsistent
\citep{Carter2019-rw}.

Hedges' \citeyearpar{hedges1992modeling} publication bias model takes
publication bias explicitly into account using a selection model,
and is arguably the most appropriate model for publication bias \citep{Carter2019-rw}.
Despite there being an $\mathtt{R}$ \citep{Team2013-tt} package
for maximum likelihood estimation of this model, called $\texttt{weightr}$
\citep{Coburn2019-ec}, the model has not yet taken off. Its maximum
likelihood estimation methods are numerically unstable and its estimates
can be off even when they converge \citep{Coburn2019-ec,Stanley2005-ng}.
The estimate of the mean effect size may be negative and of unrealistically
large magnitude, and the estimated heterogeneity parameter might be
improbably large. It turns out there are ridges in the likelihood
function that can be linked to this behavior \citep{McShane2016-rb},
but it has not been stated in clear terms exactly what the consequences
are for inferential procedures. The purpose of this note is to show
there is no confidence set for the mean effect size that has infinite
diameter with probability $0$. In the terminology of \citet{berger1999integrated},
Hedges' publication bias model belongs to the \emph{Gleser--Hwang}
class.

@@

We will assume that the random effects meta-analysis model is true
in the absence of publication bias. The mechanisms that cause publication
bias modify the density in a suitable way. Consider the case when
only significant studies at some specified level $\alpha$ are published.
Assuming one-sided tests, the \emph{p}-values are $u_{i}=\Phi(-x_{i}/\sigma_{i})$,
or normal one-sided \emph{p}-values. We will only deal with with one-sided
\emph{p}-values in this paper, as there is usually, but not always,
just one direction that is interesting to researchers, reviewers,
and editors. A one-sided \emph{p}-value can also be used if the researchers
reported a two-sided value, since $p=0.05$ for a two-sided hypothesis
corresponds to $p=0.025$ for a one-sided hypothesis, et cetera.

Define $c_{\alpha}=\Phi^{-1}(1-\alpha)$, the cutoff for significance
at level $\alpha$. The \emph{basic publication bias model} is a truncated
normal model with density
\begin{equation}
f(x_{i};\theta_{0},\sigma_{i})=\Phi\left(\frac{\theta_{0}-c_{\alpha}}{\sqrt{\sigma_{i}^{2}+\tau^{2}}}\right)^{-1}\phi(x;\theta_{0},\sqrt{\sigma_{i}^{2}+\tau^{2}})1[x_{i}/\sigma_{i}>c_{\alpha}],\label{eq:selection for significance model}
\end{equation}
This model for publication bias was introduced by \citet{hedges1984estimation}
in the context of $F$-distributions.

The basic publication bias model is unrealistic. It requires that
no non-significant studies are published. But even in the fields most
severely affected by publication bias, such as psychology, a non-negligible
number of non-significant studies are published \citep{Motyl2017-dx}.
Moreover, the basic publication bias model does not allow for different
cutoffs for significance. For instance, it is likely that some editors
will accept studies reaching a significance at $\alpha=0.025$, corresponding
to $x_{i}/\sigma_{i}>1.96$ but not at $\alpha=0.05$, corresponding
to $x_{i}/\sigma_{i}>1.64$.

These problems can be rectified by adopting the selection model for
publication bias of \citet{iyengar1988selection}, which models the
following scenario.
\begin{quote}
\textbf{Publication bias scenario.} Alice the editor receives a study
with the \emph{p}-value $u$. Her publication decision is a random
function of this \emph{p}-value. That is, she will publish the study
with some probability $w(u)$ and reject it with probability $1-w(u)$.
Every study you will ever read in Alice's journal has survived this
selection mechanism, the rest are lost forever.
\end{quote}
Let $w(u_{i})$ be a function of the \emph{p}-value $u_{i}=1-\Phi(-x_{i}/\sigma_{i})$
taking values in $[0,1]$. Then $w(u_{i})$ is a probability for every
$u_{i}$, and the selection model

@@

\citet{hedges1992modeling} studies the selection model when $w$
is a step function with fixed steps. Let $\alpha$ be a vector with
elements $0=\alpha_{0}<\alpha_{1}<\ldots<\alpha_{K}=1$ and $\rho$
be a $K$-ary non-negative, non-increasing vector having its first
element equal to $\rho_{1}=1$. Define the step function $w$ based
on $\alpha$ and $\rho$ as
\begin{equation}
w(u;\rho,\alpha)=\sum_{k=1}^{K}\rho^{k}1_{(\alpha_{k-1},\alpha_{k}]}(u).\label{eq:step function}
\end{equation}
We call the selection model with a step function \emph{Hedges' publication
bias model. }Its density is
\begin{equation}
f(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})\propto\sum_{k=1}^{K}\rho^{k}\phi(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})1_{(\alpha_{k-1},\alpha_{k}]}(u_{i}).\label{eq:hedges model}
\end{equation}
Interpreting Hedges' publication bias model is easy. When the editor
receives a study with \emph{p}-value $u$, she finds the $k$ such
that $u\in(\alpha_{k-1},\alpha_{k}]$ and accepts with probability
$\rho^{k}$. Since $\rho^{1}=1$, she always accepts when $u\in[0,\alpha_{1}]$.
The vector $\rho$ is non-increasing since a publication decision
based solely on \emph{p}-values should always act favorably towards
lower \emph{p}-values. The parameters $(\theta_{0},\tau,\rho)$ of
the model are identified when $\alpha$ is fixed \citep[Web Appendix A]{moss2019modelling}.

@@

We can write Hedges' model as a mixture model on the form
\begin{equation}
f(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})=\sum_{k=1}^{K}\pi^{k}f^{k}(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}}).\label{eq:mixture model formulation}
\end{equation}
Here $f^{K}$ is a normal density, $f^{k},k<K$ are normal densities
truncated to $[\Phi^{-1}(1-\alpha_{j-1}),\Phi^{-1}(1-\alpha))$, and
$\pi^{k}$ are mixture probabilities, i.e., $\pi^{k}>0$ for each
$k$ and $\sum_{k=1}^{K}\pi^{k}=1$. The mixture probabilities $\pi^{k}$
are functions of $(\theta_{0},\tau,\sigma_{i},\rho)$, see the appendix
(p. \pageref{eq:pi_i formula}) for their formula.

The main benefit of Hedges' publication bias model (\ref{eq:hedges model})
is how it models \emph{p}-values based publication bias directly,
there is no approximation involved. If you believe in the random effects
meta-analysis model and the \emph{p}-value based publication bias
scenario, Hedges' publication bias model is simply the correct model.
Most statistical methods correcting for publication bias in the literature
either do not make use of an explicit statistical model or do not
estimate the parameters $\theta_{0}$ and $\tau$. For instance, the
funnel plot of \citet{Egger1997-rd} is a graphical method, while
the trim-and-fill method of \citet{Duval2000-ct} is a non-parametric
method based on making the funnel plot symmetric. \citet{Stanley2005-ng,stanley2014meta}
discuss various misspecified regression-based estimators of the corrected
effect size $\theta$ based on the fixed effect Hedges' publication
bias model. The estimating \emph{p}-curve method of \citet{Simonsohn2014-cn}
and \emph{p}-uniform of \citet{Van_Assen2015-qs,Van_Aert2016-cu}
are two methods for dealing with publication bias hailing from psychology.
Both are based on a variant of the basic publication bias model, but
with fixed instead of random effects, and both employ somewhat unusual
estimation methods \citep{McShane2016-rb}. Since there is ample evidence
of heterogeneity in meta-analysis, restricting oneself to the fixed
effects meta-analysis is a mistake.

Hedges' model has some downsides. It models only bias due to selection
of \emph{p}-values, not every source of bias, such as language bias
\citep{Egger1998-kj}. Second, it may not be best model for biases
with other causes than the publication process, such as \emph{p}-hacking
\citep{simmons2011false}. \citet{moss2019modelling} propose a related
model that may be more successful at correcting for \emph{p}-hacking.

@@

Let $f^{1},f^{2},\ldots,f^{K}$ be a sequence of densities, $\pi=(\pi^{1},\pi^{2},\ldots,\pi^{K})$
a probability vector, and $p=\sum_{k\leq K}\pi^{k}f^{k}$ a mixture
distribution. We will assume that the size of $[p]$ equals the size
of any of its mixture components $[f^{k}]$ for some size $\left\Vert \cdot\right\Vert $,
i.e., $\left\Vert [p]\right\Vert =\left\Vert [f^{k}]\right\Vert $
for all $k$. Why we do this will be clear in the proof of Theorem
\ref{thm:general publication bias}, but think of it this way: If
all of $p$s mixture components have the same mean, the mean of $p$
equals the mean of any $f^{k}$. 
\begin{lem}
\label{prop:Mixture model corollary} Let $\mathcal{P}$ be a class
of $K$-ary mixture distributions and $||[p]||$ be the assumption
above. Assume there is a sequence $p_{n}=\sum_{k\leq K}\pi_{n}^{k}f_{n}^{k}$
and a subset $K'$ such that 
\begin{enumerate}
\item For all $k\in K'$, there is a density $f^{k\star}$ such that $f_{n}^{k}$
converges to $f^{k\star}$ pointwise.
\item For all mixtures $p$, $\supp p\supseteq\supp f^{k\star}$ for some
$k\in K'$.
\item For all $k\in K'$, the size of $[f_{n}^{k}]$ goes to infinity, $\left\Vert [f_{n}^{k}]\right\Vert \to\infty$.
\item The density concentrates on the components indexed by $K'$, $\lim_{n\to\infty}\sum_{k\in K'}\pi_{n}^{k}=1$.
\end{enumerate}
Then every confidence set with level $\alpha>0$ has infinite diameter
with positive probability.
\end{lem}

\begin{proof}
We employ Theorem \ref{thm:Gleser--Hwang}. By (i) and (iv), $p_{n}$
converges pointwise to the density $\sum_{k\in K'}\pi_{n}^{k\star}f_{n}^{k\star}=p^{\star}$
. That $\supp p\supseteq p^{\star}$ follows from (ii) and (iv). Finally,
from the assumption that$\left\Vert [p_{n}]\right\Vert =\left\Vert [f_{n}^{k}]\right\Vert $,
we get that $\left\Vert [p_{n}]\right\Vert \to\infty$ too. 
\end{proof}
\begin{thm}
\label{thm:general publication bias} Assume $n$ independent samples
from the publication bias model (\ref{eq:selection for significance model}),
where the selection probability $\rho$ is unknown and $\alpha$ is
known. Then any confidence set for $\theta_{0},\tau$, or $(\theta_{0},\tau)$
will have infinite diameter with positive probability.
\end{thm}

>>>>

A meta-analysis is a statistical analysis that quantitatively combines
results from separate scientific studies. When the studies measure
the same phenomenon, pooling of information allows us to predict the
common effect size with larger precision than we could have done with
one study alone. Meta-analyses are ubiquitous in the empirical sciences
and forms a key component of most systematic reviews such as Cochrane
reviews \citep{Higgins2019-vv}.

Most meta-analytic techniques assume honest and unbiased reporting
of results. But there is ample evidence that the scientific literature
is not unbiased, as studies with\emph{ }significant \textit{p}-values
tend to be published with a greater probability than other studies
\citep{Easterbrook1991-ph}, a phenomenon called publication bias
by \citet{sterling1959publication} and the file-drawer problem by
\citet{Rosenthal1979-pm}. When publication bias is present, there
is no reason to trust the results of meta-analytic methods that do
not account for it, as the parameter estimates will be inconsistent
\citep{Carter2019-rw}.

Hedges' \citeyearpar{hedges1992modeling} publication bias model takes publication bias explicitly into account using a selection model, and is arguably the most appropriate model for publication bias \citep{Carter2019-rw}. Despite there being an $\mathtt{R}$ \citep{Team2013-tt} package for maximum likelihood estimation of this model, called $\texttt{weightr}$ \citep{Coburn2019-ec}, the model has not yet taken off. Its maximum likelihood estimation methods are numerically unstable and its estimates can be off even when they converge \citep{Coburn2019-ec,Stanley2005-ng}. The estimate of the mean effect size may be negative and of unrealistically large magnitude, and the estimated heterogeneity parameter might be improbably large. It turns out there are ridges in the likelihood function that can be linked to this behavior \citep{McShane2016-rb}, but it has not been stated in clear terms exactly what the consequences are for inferential procedures. The purpose of this note is to explain why Hedges' publication bias performs poorly, by showing there is no confidence set for the mean effect size that has infinite diameter with probability $0$. %In the terminology of \citet{berger1999integrated}, Hedges' publication bias model belongs to the \emph{Gleser--Hwang} class.

@@

We will assume that the random effects meta-analysis model is true in the absence of publication bias. The mechanisms that cause publication bias modify the density in a suitable way. Consider the case when only significant studies at some specified level $\alpha$ are published. Assuming one-sided tests, the \textit{p}-values are $u_{i}=\Phi(-x_{i}/\sigma_{i})$, or normal one-sided \textit{p}-values. We will only deal with with one-sided \textit{p}-values in this paper, as there is usually, but not always, just one direction that is interesting to researchers, reviewers, and editors. A one-sided \textit{p}-value can also be used if the researchers reported a two-sided value, since $p=0.05$ for a two-sided hypothesis corresponds to $p=0.025$ for a one-sided hypothesis, et cetera.

Define $c_{\alpha}=\Phi^{-1}(1-\alpha)$, the cutoff for significance at level $\alpha$. The \emph{basic publication bias model} is a truncated normal model with density
\begin{equation}
f(x_{i};\theta_{0},\sigma_{i})=\Phi\left(\frac{\theta_{0}-c_{\alpha}}{\sqrt{\sigma_{i}^{2}+\tau^{2}}}\right)^{-1}\phi(x;\theta_{0},\sqrt{\sigma_{i}^{2}+\tau^{2}})1[x_{i}/\sigma_{i}>c_{\alpha}],\label{eq:selection for significance model}
\end{equation}
where $1[A]$ is the characteristic function of $A$. This model for publication bias was introduced by \citet{hedges1984estimation} in the context of $F$-distributions.

The basic publication bias model is unrealistic. It requires that no non-significant studies are published. But even in the fields most severely affected by publication bias, such as psychology, a non-negligible number of non-significant studies are published \citep{Motyl2017-dx}. Moreover, the basic publication bias model does not allow for different cutoffs for significance. It is likely that some editors will accept studies reaching a significance at $\alpha=0.025$, corresponding to $x_{i}/\sigma_{i}>1.96$ but not at $\alpha=0.05$, corresponding to $x_{i}/\sigma_{i}>1.64$.

These problems can be rectified by adopting the selection model for publication bias of \citet{iyengar1988selection}, which models the following scenario.
\begin{quote}
\textbf{Publication bias scenario.} Alice the editor receives a study with the \textit{p}-value $u$. Her publication decision is a random function of this \textit{p}-value. That is, she will publish the study with some probability $w(u)$ and reject it with probability $1-w(u)$. Every study you will ever read in Alice's journal has survived this selection mechanism, the rest are lost forever.
\end{quote}
Let $w(u_{i})$ be a function of the \textit{p}-value $u_{i}=\Phi(-x_{i}/\sigma_{i})$ taking values in $[0,1]$. Then $w(u_{i})$ is a probability for every $u_{i}$, and the selection model

@@

\citet{hedges1992modeling} studies the selection model when $w$ is a step function with fixed steps. Let $\alpha$ be a vector with elements $0=\alpha_{0}<\alpha_{1}<\ldots<\alpha_{K}=1$ and $\rho$ be a $K$-ary non-negative, non-increasing vector having its first element equal to $\rho_{1}=1$ for identifiability. Define the step function $w$ based on $\alpha$ and $\rho$ as
\begin{equation}
w(u;\rho,\alpha)=\sum_{k=1}^{K}\rho^{k}1_{(\alpha_{k-1},\alpha_{k}]}(u).\label{eq:step function}
\end{equation}
We call the selection model with a step function \emph{Hedges' publication bias model}. Its density is
\begin{equation}
f(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})\propto\sum_{k=1}^{K}\rho^{k}\phi(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})1{(\alpha_{k-1},\alpha_{k}]}(u_{i}).\label{eq:hedges model}
\end{equation}
Interpreting Hedges' publication bias model is easy. When the editor receives a study with \textit{p}-value $u$, she finds the $k$ such that $u\in(\alpha_{k-1},\alpha_{k}]$ and accepts with probability $\rho^{k}$. Since $\rho^{1}=1$, she always accepts when $u\in[0,\alpha_{1}]$. The vector $\rho$ is non-increasing since a publication decision based solely on \textit{p}-values should always act favorably towards lower \textit{p}-values. The parameters $(\theta_{0},\tau,\rho)$ of the model are identified when $\alpha$ is fixed \citep[Web Appendix A]{moss2019modelling}.

@@

We can write Hedges' model as a mixture model on the form
\begin{equation}
f(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}})=\sum_{k=1}^{K}\pi^{k}f^{k}(x_{i};\theta_{0},\sqrt{\tau^{2}+\sigma_{i}^{2}}).\label{eq:mixture model formulation}
\end{equation}
Here $f^{K}$ is a normal density, $f^{k},k<K$ are normal densities
truncated to $[\Phi^{-1}(1-\alpha_{k-1}),\Phi^{-1}(1-\alpha_k))$, and $\pi^{k}$ are mixture probabilities, i.e., $\pi^{k}>0$ for each $k$ and $\sum_{k=1}^{K}\pi^{k}=1$. The mixture probabilities $\pi^{k}$ are functions of $(\theta_{0},\tau,\sigma_{i},\rho)$, see the appendix (p. \pageref{eq:pi_i formula}) for their formula.

The main benefit of Hedges' publication bias model (\ref{eq:hedges model}) is how it models \textit{p}-values based publication bias directly, there is no approximation involved. If you believe in the random effects meta-analysis model and the \textit{p}-value based publication bias scenario, Hedges' publication bias model is simply the correct model. Most statistical methods correcting for publication bias in the literature either do not make use of an explicit statistical model or do not estimate the parameters $\theta_{0}$ and $\tau$. For instance, the funnel plot of \citet{Egger1997-rd} is a graphical method, while the trim-and-fill method of \citet{Duval2000-ct} is a non-parametric method based on making the funnel plot symmetric. \citet{Stanley2005-ng,stanley2014meta} discuss various misspecified regression-based estimators of the corrected effect size $\theta$ based on the fixed effect Hedges' publication bias model. The estimating \textit{p}-curve method of \citet{Simonsohn2014-cn} and \textit{p}-uniform of \citet{Van_Assen2015-qs,Van_Aert2016-cu} are two methods for dealing with publication bias hailing from psychology. Both are based on a variant of the basic publication bias model, but with fixed instead of random effects, and both employ somewhat unusual estimation methods \citep{McShane2016-rb}. Since there is ample evidence of heterogeneity in meta-analysis, restricting oneself to the fixed effects meta-analysis is a mistake.

Hedges' model has some downsides. It models only bias due to selection
of \textit{p}-values, not every source of bias, such as language bias
\citep{Egger1998-kj}. Second, it may not be best model for biases
with other causes than the publication process, such as \textit{p}-hacking
\citep{simmons2011false}. \citet{moss2019modelling} propose a related
model that may be more successful at correcting for \textit{p}-hacking.

@@

Let $f^{1},f^{2},\ldots,f^{K}$ be a sequence of densities, $\pi=(\pi^{1},\pi^{2},\ldots,\pi^{K})$
a probability vector, and $p=\sum_{k\leq K}\pi^{k}f^{k}$ a mixture
distribution. We will assume that the size of $[p]$ equals the size
of any of its mixture components $[f^{k}]$ for some size $\left\Vert \cdot\right\Vert $,
i.e., $\left\Vert [p]\right\Vert =\left\Vert [f^{k}]\right\Vert $
for all $k$. Why we do this will be clear in the proof of Theorem
\ref{thm:general publication bias}, but think of it this way: If
all of $p$s mixture components have the same mean, the mean of $p$
equals the mean of any $f^{k}$. 
\begin{lem}
\label{prop:Mixture model corollary} Let $\mathcal{P}$ be a class
of $K$-ary mixture distributions and $||[p]||$ be as assumed above. Assume there is a sequence $p_{n}=\sum_{k\leq K}\pi_{n}^{k}f_{n}^{k}$
and a subset $K'$ such that 
\begin{enumerate}
\item For all $k\in K'$, there is a density $f^{k\star}$ such that $f_{n}^{k}$
converges to $f^{k\star}$ pointwise.
\item For all mixtures $p$, $\supp p\supseteq\supp f^{k\star}$ for some
$k\in K'$.
\item For all $k\in K'$, the size of $[f_{n}^{k}]$ goes to infinity, $\left\Vert [f_{n}^{k}]\right\Vert \to\infty$.
\item The density concentrates on the components indexed by $K'$, $\lim_{n\to\infty}\sum_{k\in K'}\pi_{n}^{k}=1$.
\end{enumerate}
Then every confidence set with level $\alpha>0$ has infinite diameter
with positive probability.
\end{lem}

\begin{proof}
We employ Theorem \ref{thm:Gleser--Hwang}. By (i) and (iv), $p_{n}$
converges pointwise to the density $$\sum_{k\in K'}\pi_{n}^{k\star}f_{n}^{k\star}=p^{\star}.$$
That $\supp p\supseteq p^{\star}$ follows from (ii) and (iv). Finally,
from the assumption that$\left\Vert [p_{n}]\right\Vert =\left\Vert [f_{n}^{k}]\right\Vert $,
we get that $\left\Vert [p_{n}]\right\Vert \to\infty$ too. 
\end{proof}
\begin{thm}
\label{thm:general publication bias} Assume $n$ independent samples
from the publication bias model (\ref{eq:selection for significance model}),
where the selection probability $\rho$ is unknown and $\alpha$ is
known. Then any confidence set for $\theta_{0},\tau$, or $(\theta_{0},\tau)$
will have infinite diameter with positive probability.
\end{thm}
